Revision of JMLR-14-133: Point-to-Point Reply

### 2014-09-07 pkg/partykit/inst/mloss/review2/p2p.tex
improvements and reply to AE
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@607 255f4249-2263-4417-b1ea-8498779e27a7

\documentclass{article}
\usepackage{a4wide,natbib}

@@

\textit{%
I regret to inform you that your submission ``partykit: A modular toolkit
for recursive partytioning in R'' has been rejected from publication at JMLR
Track for Machine Learning Open Source Software.  Two of the reviewers of
the previous version were enlisted for this version, and one new reviewer
was added.  While the reviewers voted for ``accept'', the actual reviews
were less positive and raised several serious concerns about the project. 
If the authors choose to resubmit the project, please include a description
addressing the reviewers' concerns.
}

@@

Thank you for the constructive reviews and feedback and for giving us the
chance to revise our manuscript/software and submit it to the MLOSS track
again. Please find attached ...

The most important changes ... discussion rewritten ...

@@

\textit{%
This is a clean and concise description of an interesting set of tools for
data analysis with recursive partitioning.  It is mainly a description or
announcement, and has minimal polemic or pedagogical value.
}

@@

\textit{%
It will be a useful reference for those a basic outline of the software's
capabilities, and provides a useful bibliography.
}

@@

\textit{%
This paper describes work that, in my understanding, does two things:
\begin{enumerate}
\item It provides infrastructure that can in principle be used with any
      recursive partitioning implementation, and that at this point 
      interfaces to several different widely used implementations, 
      additional to those directly available in R's partykit package.
\item It provides implementations of two specialized tree-based methods.
\end{enumerate}
}

@@

When we started working on \emph{partykit}, we redesigned all classes from the
earlier \emph{party} package from scratch and also reimplemented \texttt{ctree} and \texttt{mob}
without relatively little overlap between the old and new code bases. As a
consequence of the much cleaner \emph{partykit} design, the code base for both
\texttt{ctree} and \texttt{mob} could be reduced considerably while at the same time
both implementations gained more flexibility and additional options by leveraging
the new class structure.

@@

\textit{%
The software has a useful role in providing commonality across diverse
recursive partitioning methods and implementations: for structuring printed
output, for generating predictions, and for plotting results.  Very
extensive documentation, some distributed with the package, is available
from references given in the paper.
}

@@

\textit{Quibbles and queries are:}
\begin{enumerate}
\item 
\textit{%
The legend for Table 1 begins: ``Selected tree algorithms than can be
interfaced through partykit...''  Surely it is the output from selected
implementations that can be interfaced, not the algorithm used.
}

@@

\textit{%
The word `algorithm' seems to me anyway inappropriate in this context. The
same (or very similar) algorithm may have different implementations.  Different
methods with their inevitably different algorithms (including, in partykit:
classical `constant' trees, model-based trees, and conditional inference
trees) further widen the range of implementations.
}

@@

\item 
\textit{%
The distinction between implementation (what the authors call `algorithm')
and output is crucial also because as far as I am able to judge, partykit's
abilities are limited to taking over and processing some part only of the
output from, e.g., R's rpart package.  As I understand, there is no ability
to use or process, e.g., results from rpart's on the fly cross-validation.
}

@@

Yes, this is correct. It is not possible to directly work on the \texttt{rpart}
results after coercing to the \texttt{constparty} class.  The reason is that not
all constant-fit trees are based on a risk or loss function and hence there
is no uniform approach for performing cost-complexity pruning.
Hence, the current implementation assumes that the pruning has already been
performed within \texttt{rpart} (or \texttt{J48}) prior to coercion.

@@

\textit{%
Would it be straightforward to incorporate structure that is able to
accommodate such information?  See also point 3 that now follows.
}

@@

\item 
\textit{%
I note the comment on the help page for the partykit function ctree() that
the inference process is based on permutation tests, with one or other type
of adjustment for multiple testing.  Then follows the strong claim that:
This statistical approach ensures that the right sized tree is grown and no
form of pruning or cross-validation or whatsoever is needed.  I think it
pertinent to ask what has been done to judge the effectiveness of this
approach, assessed using separate test data, against the approach used in,
e.g., rpart or CART?  This is an issue for the help page, rather than
perhaps for the paper.
}

@@

\item 
\textit{%
How well-adapted are the tools in partykit for implementing random forest
or other ensemble type approaches? I note that partykit does not (yet?) have
an equivalent of the cforest() function that is in the older party package.
}

@@

\textit{%
At the very least, I would like to see brief comments on where splitting and
accuracy assessment strategies, short of use of separate test data, fit in
or around the toolkit.
}

@@

\textit{To what extent are toolkit functions independent of
the methodology used for the partykit functions ctree() and mob()?}

@@

\textit{If necessary, some of the historical detail in the overview might be replaced
by references to discussion that appears elsewhere, in order to make way for
comment on these points.}

@@

\textit{%
The paper describes an R library, `partykit', that appears to be a unified
way of representing decision tree structures (or recursive partitioning
structures) in R and manipulating or visualizing them.  The authors note
that their library contains support for representing the output of trees
created with other packages, and the paper appears to function as a very
basic tutorial with links to very detailed R vignettes.  The package also
implements a few of the authors' own recursive partitioning algorithms.}

@@

\textit{%
The project seems to have fairly minor activity, and appears to have had a
Google Summer of Code project in previous years.  The subversion repository
seems to indicate that the code is still being (somewhat) actively developed
despite a relative lack of commits between 2009 and 2013.
}

@@

However, during 2009 and 2013 the package was actively maintained with six
CRAN releases, e.g., fixing bugs or incorporating smaller user requests. Also
during this time, two offspring packages (\emph{evtree} and \emph{psychotree}) were
developed that leverage the \emph{partykit} infrastructure. These brought
only some smaller design or implementation deficits to light that did not
necessitate (re-)writing/modifying large parts of the code.

@@

\textit{%
There appears to be no bug tracker or open mailing list.
}

@@

E-mails to the maintainer address were always possible and this is still a very common
way of communicating with CRAN package maintainers (and developers). In the past,
this means of communication with us was used fairly regularly by both end-users
and other package developers.

As \textsf{R}-Forge easily allows for having mailing lists, forums, etc., we
have added a discussion forum and also point the readers to this in the rewritten
Section~5 of our manuscript.

@@

\textit{%
There is knowledge of the package in the R ecosystem, but partykit does not seem
to be as popular as other R packages (admittedly, partykit is a niche product).
}

@@

As of now, CRAN package web page lists seven reverse dependencies (in two of
which we have no involvement at all and two others that are not maintained by us).
Only 6\% of all CRAN packages have more reverse dependencies than that. And we hope
that some of the 24~reverse dependencies of \emph{party} will switch to \emph{partykit}
(only 2\% of all CRAN packages have more than 24 reverse dependencies) now that the
recommended reference implementations of \texttt{ctree} and \texttt{mob} are in \emph{partykit}.
Moreover, according to Google Scholar the \texttt{ctree} and \texttt{mob} papers
published in \emph{Journal of Computational and Graphical Statistics} were cited
643 and 103 times, respectively.

@@

\textit{%
Support for trees produced by other libraries is provided, but I am not
certain how appealing the infrastructure will be to developers who want to
represent trees from other systems.  It seems like for non-PMML formatted
trees, implementing a conversion layer to partykit objects wouldn't be any
easier or harder than other packages. Clearly, for developing new tree
algorithms, though, partykit would be a good choice.
}

@@

Our experience is not that any tree model can be converted in the same easy
or hard way to any other packages.

While implementing the original code for \texttt{ctree} 10~years ago, we could
not easily leverage infrastructure that existed at the time because it had way too many
algorithm-specific things hardwired into it. While \texttt{rpart}
always allowed to change risk functions, doing something outside of the
CART-way of finding splits was not really possible. Also printing and plotting
facilities were somewhat limited and not directly able to directly include $p$-values
in the inner nodes, for example.
Our \emph{party} implementation of \texttt{ctree} then made the same mistake of
wiring too many algorithm-specific requirements into the class. Hence, it really
had to be abused to capture \texttt{mob} trees. The multi-way splits of
\texttt{J48} from \emph{RWeka} could not be captured by either \emph{rpart}
or \emph{party}.

@@

\textit{%
Section 5 is very qualitative and vague.  The authors should definitely
improve this in future revisions.  It is written, "lean data structures are
employed", "the `party' structures are very lean", and also "The recursive
node structure itself is extremely lightweight" but without any sort of
quantitative details, it's difficult to believe those statements without
diving into the code.
}

@@

\textit{%
In my opinion the authors could do one of these things to validate their
claims: provide proof that the data structures are minimal (with respect to
other implementations), show that very large tree structures can be easily
represented and handled by partykit, or simply leave out or significantly
shorten the discussion of speed and memory usage entirely.
}

@@

We have added a demo to the package that briefly illustrates the
memory and speed considerations that we alluded to in the initial submission.
This can be run interactively in \textsf{R} via
\texttt{demo("memory-speed", package = "partykit")}. However, it is not
mentioned in the manuscript as this is not the main focus of the package.

@@

\textit{%
Overall, partykit is a well-designed and extensively documented package, but
I have some reservations about the project's activity level and popularity
going into the future.  The developers should consider setting up a mailing
list or bug tracker to make the project more open to new contributors.
}

>>>>

\documentclass{article}
\usepackage{a4wide}
\usepackage[authoryear,round]{natbib}

@@

\textbf{\textit{%
I regret to inform you that your submission ``partykit: A modular toolkit
for recursive partytioning in R'' has been rejected from publication at JMLR
Track for Machine Learning Open Source Software.  Two of the reviewers of
the previous version were enlisted for this version, and one new reviewer
was added.  While the reviewers voted for ``accept'', the actual reviews
were less positive and raised several serious concerns about the project. 
If the authors choose to resubmit the project, please include a description
addressing the reviewers' concerns.
}}

@@

Thank you for the constructive reviews and feedback and for giving us the
chance to revise our manuscript/software and submit it to the MLOSS track
again. Please find attached the revision of our four-page manuscript
and a new version (0.8-1) of the \emph{partykit} package which is already
available from \texttt{http://CRAN.R-project.org/package=partykit} as well.

The most important changes are a completely rewritten Section~5 in the four-page
manuscript (Discussion and Outlook) and an extended Section~3 in the \texttt{constparty}
vignette (showing how new tree algorithms can be implemented using \emph{partykit}).
Moreover, an online support forum for \emph{partykit} has been established
on \textsf{R}-Forge at \texttt{https://R-Forge.R-project.org/forum/forum.php?forum\_id=852}.
Further details of the revision are discussed in the point-to-point reply
to the reviewers.

Note that the revised four-page manuscript now has the acknowledgments (four lines)
on the fifth page along with the references. We are not sure whether this
conforms with the MLOSS style guidelines: if not, we will, of course, streamline
the main manuscript to place the acknowledgments on the first four pages as well.

@@

\textbf{\textit{%
This is a clean and concise description of an interesting set of tools for
data analysis with recursive partitioning.  It is mainly a description or
announcement, and has minimal polemic or pedagogical value.
}}

@@

\textbf{\textit{%
It will be a useful reference for those a basic outline of the software's
capabilities, and provides a useful bibliography.
}}

@@

\textbf{\textit{%
This paper describes work that, in my understanding, does two things:
\begin{enumerate}
\item It provides infrastructure that can in principle be used with any
      recursive partitioning implementation, and that at this point 
      interfaces to several different widely used implementations, 
      additional to those directly available in R's partykit package.
\item It provides implementations of two specialized tree-based methods.
\end{enumerate}
}}

@@

When we started working on \emph{partykit}, we redesigned all classes from the
earlier \emph{party} package from scratch and also reimplemented \texttt{ctree} and \texttt{mob}
with relatively little overlap between the old and new code bases. As a
consequence of the much cleaner \emph{partykit} design, the code base for both
\texttt{ctree} and \texttt{mob} could be reduced considerably while at the same time
both implementations gained more flexibility and additional options by leveraging
the new class structure.

@@

\textbf{\textit{%
The software has a useful role in providing commonality across diverse
recursive partitioning methods and implementations: for structuring printed
output, for generating predictions, and for plotting results.  Very
extensive documentation, some distributed with the package, is available
from references given in the paper.
}}

@@

\textbf{\textit{%
Quibbles and queries are:
}}
\begin{enumerate}
\item 
\textbf{\textit{%
The legend for Table 1 begins: ``Selected tree algorithms than can be
interfaced through partykit...''  Surely it is the output from selected
implementations that can be interfaced, not the algorithm used.
}}

@@

\textbf{\textit{%
The word `algorithm' seems to me anyway inappropriate in this context. The
same (or very similar) algorithm may have different implementations.  Different
methods with their inevitably different algorithms (including, in partykit:
classical `constant' trees, model-based trees, and conditional inference
trees) further widen the range of implementations.
}}

@@

\item 
\textbf{\textit{%
The distinction between implementation (what the authors call `algorithm')
and output is crucial also because as far as I am able to judge, partykit's
abilities are limited to taking over and processing some part only of the
output from, e.g., R's rpart package.  As I understand, there is no ability
to use or process, e.g., results from rpart's on the fly cross-validation.
}}

@@

Yes, this is correct. It is not possible to directly work on the \texttt{rpart}
results after coercion to the \texttt{constparty} class.  The reason is that not
all constant-fit trees are based on a risk or loss function and hence there
is no uniform approach for performing cost-complexity pruning.
Therefore, the current implementation assumes that the pruning has already been
performed within \texttt{rpart} (or \texttt{J48}) prior to coercion.

@@

\textbf{\textit{%
Would it be straightforward to incorporate structure that is able to
accommodate such information?  See also point 3 that now follows.
}}

@@

\item 
\textbf{\textit{%
I note the comment on the help page for the partykit function ctree() that
the inference process is based on permutation tests, with one or other type
of adjustment for multiple testing.  Then follows the strong claim that:
This statistical approach ensures that the right sized tree is grown and no
form of pruning or cross-validation or whatsoever is needed.  I think it
pertinent to ask what has been done to judge the effectiveness of this
approach, assessed using separate test data, against the approach used in,
e.g., rpart or CART?  This is an issue for the help page, rather than
perhaps for the paper.
}}

@@

\item 
\textbf{\textit{%
How well-adapted are the tools in partykit for implementing random forest
or other ensemble type approaches? I note that partykit does not (yet?) have
an equivalent of the cforest() function that is in the older party package.
}}

@@

\textbf{\textit{%
At the very least, I would like to see brief comments on where splitting and
accuracy assessment strategies, short of use of separate test data, fit in
or around the toolkit.
}}

@@

\textbf{\textit{%
To what extent are toolkit functions independent of
the methodology used for the partykit functions ctree() and mob()?
}}

@@

\textbf{\textit{%
If necessary, some of the historical detail in the overview might be replaced
by references to discussion that appears elsewhere, in order to make way for
comment on these points.
}}

@@

\textbf{\textit{%
The paper describes an R library, `partykit', that appears to be a unified
way of representing decision tree structures (or recursive partitioning
structures) in R and manipulating or visualizing them.  The authors note
that their library contains support for representing the output of trees
created with other packages, and the paper appears to function as a very
basic tutorial with links to very detailed R vignettes.  The package also
implements a few of the authors' own recursive partitioning algorithms.
}}

@@

\textbf{\textit{%
The project seems to have fairly minor activity, and appears to have had a
Google Summer of Code project in previous years.  The subversion repository
seems to indicate that the code is still being (somewhat) actively developed
despite a relative lack of commits between 2009 and 2013.
}}

@@

However, during 2009 and 2013 the package was actively maintained with six
CRAN releases, e.g., fixing bugs or incorporating smaller user requests. Also
during this time, two offspring packages (\emph{evtree} and \emph{psychotree}) were
developed that leverage the \emph{partykit} infrastructure. These brought
only some smaller design or implementation deficits to light that did not
necessitate (re-)writing/modifying large parts of the \emph{partykit} code.

@@

\textbf{\textit{%
There appears to be no bug tracker or open mailing list.
}}

@@

E-mails to the maintainer address were always possible and this is a rather common
way of communicating with CRAN package maintainers (and developers). In the past,
this means of communication with us was used fairly regularly by both end-users
and other package developers.

As \textsf{R}-Forge easily allows for having mailing lists, forums, etc., we
have added a discussion forum (\texttt{https://R-Forge.R-project.org/forum/forum.php?forum\_id=852})
and also point the readers to this in the rewritten Section~5 of our manuscript.

@@

\textbf{\textit{%
There is knowledge of the package in the R ecosystem, but partykit does not seem
to be as popular as other R packages (admittedly, partykit is a niche product).
}}

@@

As of now, the CRAN package web page lists seven reverse dependencies (in two of
which we have no involvement at all and two others that are not maintained by us).
Only 6\% of all CRAN packages have more reverse dependencies than that. And we hope
that some of the 24~reverse dependencies of \emph{party} will switch to \emph{partykit}
(only 2\% of all CRAN packages have more than 24 reverse dependencies) now that the
recommended reference implementations of \texttt{ctree} and \texttt{mob} are in \emph{partykit}.
Moreover, according to Google Scholar the \texttt{ctree} and \texttt{mob} papers
published in \emph{Journal of Computational and Graphical Statistics} were cited
643 and 103 times, respectively.

@@

\textbf{\textit{%
Support for trees produced by other libraries is provided, but I am not
certain how appealing the infrastructure will be to developers who want to
represent trees from other systems.  It seems like for non-PMML formatted
trees, implementing a conversion layer to partykit objects wouldn't be any
easier or harder than other packages. Clearly, for developing new tree
algorithms, though, partykit would be a good choice.
}}

@@

Our experience is not that any tree model can be converted in the same easy
or hard way to any other package.

While implementing the original code for \texttt{ctree} 10~years ago, we could
not easily leverage infrastructure that existed at the time because it had way too many
algorithm-specific things hardwired into it. While \texttt{rpart}
always allowed to change risk functions, doing something outside of the
CART-way of finding splits was not really possible. Also printing and plotting
facilities were somewhat limited and not able to directly include $p$-values
in the inner nodes, for example.
Our \emph{party} implementation of \texttt{ctree} then made the same mistake of
wiring too many algorithm-specific requirements into the class. Hence, it really
had to be somewhat abused to capture \texttt{mob} trees. And the multi-way splits of
\texttt{J48} from \emph{RWeka} could not be captured by either \emph{rpart}
or \emph{party}.

@@

\textbf{\textit{%
Section 5 is very qualitative and vague.  The authors should definitely
improve this in future revisions.  It is written, ``lean data structures are
employed", ``the `party' structures are very lean", and also ``The recursive
node structure itself is extremely lightweight" but without any sort of
quantitative details, it's difficult to believe those statements without
diving into the code.
}}

@@

\textbf{\textit{%
In my opinion the authors could do one of these things to validate their
claims: provide proof that the data structures are minimal (with respect to
other implementations), show that very large tree structures can be easily
represented and handled by partykit, or simply leave out or significantly
shorten the discussion of speed and memory usage entirely.
}}

@@

We have added a demo to the package that briefly illustrates the
memory and speed considerations that we alluded to in the initial submission
by comparison to \texttt{rpart} and \texttt{J48}.
This can be run interactively in \textsf{R} via
\texttt{demo("memory-speed", package = "partykit")}. However, it is not
mentioned in the manuscript as this is not the main focus of the package.

@@

\textbf{\textit{%
Overall, partykit is a well-designed and extensively documented package, but
I have some reservations about the project's activity level and popularity
going into the future.  The developers should consider setting up a mailing
list or bug tracker to make the project more open to new contributors.
}}
